---
title: "Merged & Cleaning"
author: "Ugochi Ugbomeh"
date: "2024-08-13"
output: html_document
---

## Merged Train Dataset (Merged_train_1.xlsx) merged_1422
This dataset contains institutional data about international PG students enrollment numbers, visa requirements, and associated costs.
It includes various fields that describe the level of study, type of visa, fees, and details specific to different countries.
##Variables:
Level Of Study: Indicates the academic level (e.g., "All postgraduate").
Mode Of Study: Specifies the mode of study (e.g., "All", "Full-time").
Academic Year: Represents the academic year in a format like "2014/2015".
Visa Code: The type of visa (e.g., "Tier 4 (General)").
Visa Fee (£): The cost of the visa in GBP.
Visa-NHS Charge (£): The additional National Health Service charge associated with the visa.
Visa After Graduation: Information on the visa status after graduation.
Visa Work Requirements After Graduation: Details about work requirements after graduation.
Salary Requirements: Indicates any salary requirements related to the visa or employment.
Country Columns (e.g., China, India, Japan, etc.): These columns represent the number of enrolled students from that country
------------------
##FX Rates Dataset (rates_semi_clean.csv) = rates_1422
This dataset provides foreign exchange (FX) rates between GBP (British Pound) and various currencies for different years and months. 
This dataset and it's variables act as a key factor in predicting international PG students enrollment numbers in scottish universities.
#Variables:
Year: The calendar year (e.g., 2014, 2015).
Month: The month of the year (e.g., 1 for January, 2 for February).
Date: The month and year in a combined format (e.g., "Dec-19").
Libya_fx,Congo_fx,Ghana_fx....: These columns represent the exchange rate of GBP to respective foreign currencies. 
Each column is named with the pattern (country_fx).
-------------------


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("viridis")
#install.packages("kableExtra")
install.packages("factoextra")
```


```{r}
# Load necessary packages
library(dplyr)
library(readr)
library(ggplot2)
library(viridis)
library(tidyverse)
library(reshape2)
library(readxl)
library(RColorBrewer)
library(dummy)
library(corrplot)
library(randomForest)
library(tseries)
library(caret)
```


```{r}
# Load the uploaded CSV file
file_path1 = 'C:/Users/gochi/Desktop/CMM-500/Project Data/merged_train_1.csv'
merged_1422 = read.csv(file_path1, stringsAsFactors = T)
file_path2 = 'C:/Users/gochi/Desktop/CMM-500/Project Data/rates_semi_clean.csv'
rates_1422 = read.csv(file_path2, stringsAsFactors = T)
file_path3 = 'C:/Users/gochi/Desktop/CMM-500/Project Data/merged_1422_result.csv'
merged_1422_result = read.csv(file_path3, stringsAsFactors = T)
```


```{r}
# Convert the Year, Month, and Date columns to characters
rates_1422$Year <- as.factor(rates_1422$Year)
rates_1422$Month <- as.factor(rates_1422$Month)
rates_1422$Date <- as.factor(rates_1422$Date)
```

```{r}
str(rates_1422)
```


#View Summary
```{r}
summary(merged_1422)
summary(rates_1422)
```

##------------------------------------------------------DATA PRE-PROCESSING and CLEANING-------------------------------------------------
##Reshaping Fx Dataset
```{r}
# Pivot to long format to create Country and FX Rate columns
# Explicitly listing each country
fx_long_1422 <- rates_1422 %>%
  pivot_longer(cols = c("Libya_fx", "Ghana_fx", "Nigeria_fx", "China_fx", "Kenya_fx", 
                        "Gambia_fx", "India_fx", "Pakistan_fx", "Turkey_fx", 
                        "Oman_fx", "Japan_fx", "Thailand_fx", "Kuwait_fx","Tanzania_fx", "Uganda_fx"),
               names_to = "Country", 
               values_to = "FX_Rate")

# Step 3: Clean up Country names (optional, e.g., remove "_fx" suffix)
fx_long_1422 <- fx_long_1422 %>%
  mutate(Country = gsub("_fx", "", Country))

fx_long_1422$Month = NULL
fx_long_1422$Date = NULL
fx_long_1422$Year = as.character(fx_long_1422$Year)

# Step 4: View the result
head(fx_long_1422)
```


##Still Reshaping ( Enrollment dataset)
```{r}
#Pivot to long format to create Country and Enrollment columns
# Explicitly listing each country
mg_long_1422 <- merged_1422 %>%
  pivot_longer(cols = c("China", "India", "Japan", "Pakistan", "Thailand", 
                        "Kuwait", "Oman",  "Turkey", "Libya", 
                        "Ghana", "Nigeria", "Kenya", "Gambia", "Tanzania", "Uganda"),
               names_to = "Country", 
               values_to = "Enrollment")

mg_long_1422 <- mg_long_1422 %>%
  mutate(Fiscal_Year_Start = as.character(sub("/.*", "", Academic.Year)),
         Fiscal_Year_End = as.character(sub(".*/", "", Academic.Year)))

mg_long_1422 <- mg_long_1422 %>%
  mutate(Fiscal_Year = as.numeric(sub("/.*", "", Academic.Year))) %>%  # Extract the start year
  select(-Fiscal_Year_Start, -Fiscal_Year_End)  # Remove unnecessary columns


# Step 3: View the result
head(mg_long_1422)

```

```{r}
#import saved merged file due to errors

```




#AGGREGATE DATA
-----
```{r}
# Step 1: Aggregate the Enrollment column, including HE.Provider in the grouping
aggregated_enrollment <- merged_1422_result %>%
  group_by(Fiscal_Year, Country, FX_Rate, HE.Provider) %>%
  summarise(Total_Enrollment = sum(Enrollment, na.rm = TRUE))

# Step 2: Merge the aggregated enrollment back into the original data
aggregated_data <- merged_1422_result %>%
  left_join(aggregated_enrollment, by = c("Fiscal_Year", "Country", "FX_Rate", "HE.Provider"))

# Remove duplicates based on the relevant columns
aggregated_data_u <- aggregated_data %>%
  distinct(HE.Provider, Fiscal_Year, Country, FX_Rate, .keep_all = TRUE)

# Remove the Total_Enrollment column if not needed
aggregated_data_u$Total_Enrollment <- NULL

# View the final data
head(aggregated_data_u)

#Save or view the result
#write.csv(aggregated_data_u, "C:/Users/gochi/Desktop/CMM-500/Project Data/aggegrated_data_u.csv", row.names = FALSE)
```




```{r}
# Check for fully duplicated rows
# Identify duplicates based on Fiscal_Year, Country, and FX_Rate
duplicates <- aggregated_data_u %>%
  group_by(Fiscal_Year, Country) %>%
  filter(n() > 1) %>%
  arrange(Fiscal_Year, Country, FX_Rate)

# View the duplicates
print(duplicates)
```


```{r}
# Assuming the dataset is in a variable called dataset

# Step 1: Aggregate FX_Rate while retaining all other columns
aggregated_data_u <- aggregated_data_u %>%
  group_by(Fiscal_Year, Country, HE.Provider, Level.Of.Study, Mode.Of..Study, 
           Academic.Year, Visa.Code, Visa.Fee...., Visa.NHS.Charge...., 
           Visa.After.Graduation, Visa.Work.Requirements.After.Graduation, 
           Salary.Requirements, Dependents.Allowed, Visa.Fee..Dependents, 
           Visa.NHS.Charge.Depen, HE.Provider.Term.1, HE.Provider.Term.2) %>%
  summarise(Enrollment = sum(Enrollment, na.rm = TRUE),
            FX_Rate = round(mean(FX_Rate, na.rm = TRUE), 2),  # Aggregate and round FX_Rate
            .groups = 'drop') %>%
  distinct()  # Remove any remaining duplicates

# Remove the Date column if it's no longer needed
#aggregated_data_u <- aggregated_data_u %>%
  #select(-Date)

# View the final data
head(aggregated_data_u)
```

```{r}
# Permanently delete rows where Mode.Of..Study is "All"
aggregated_data_u <- aggregated_data_u %>%
  filter(Mode.Of..Study != "All")

# View the updated dataset
head(aggregated_data_u)
aggregated_data_u$Fiscal_Year =as.character(aggregated_data_u$Fiscal_Year)

```

#Merging datasets ( Enrolment + FX)
```{r}
# Create a Fiscal Year column by extracting start and end years from "Academic Year"
# Expand the dataset to have a row for each fiscal year
mg_long_1422 <- mg_long_1422 %>%
  pivot_longer(cols = c(Fiscal_Year), 
               names_to = "Type", 
               values_to = "Fiscal_Year") %>%
  select(-Type)  # Remove the Type column as it's no longer needed

# Ensure the Year column in fx_long_1422 is numeric

fx_long_1422 <- fx_long_1422 %>%
  mutate(Year = as.numeric(as.character(Year)))

# Merge with the FX rates dataset using both "Fiscal_Year" and "Country"
merged_1422_result <- merge(mg_long_1422, fx_long_1422, 
                            by.x = c("Fiscal_Year", "Country"), 
                            by.y = c("Year", "Country"), 
                            all.x = TRUE)

# Clean up the merged dataset
merged_1422_result <- merged_1422_result %>%
  distinct()  # Remove any potential duplicates

# View the result
head(merged_1422_result)

summary(merged_1422_result)
```


```{r}
merged_1422_result$Fiscal_Year = as.factor(merged_1422_result$Fiscal_Year)
merged_1422_result$Visa.Fee....= as.factor(merged_1422_result$Visa.Fee...)
merged_1422_result$Visa.NHS.Charge....= as.factor(merged_1422_result$Visa.NHS.Charge....)
merged_1422_result$Visa.Fee..Dependents = as.factor(merged_1422_result$Visa.Fee..Dependents)
merged_1422_result$Visa.NHS.Charge.Depen = as.factor(merged_1422_result$Visa.NHS.Charge.Depen)
summary(merged_1422_result)
#Save or view the result
#write.csv(merged_1422_result, "C:/Users/gochi/Desktop/CMM-500/Project Data/merged_1422_result.csv", row.names = FALSE)
```

# Still pre-processing
----------------------------------------
multiple rows have been duplicated based on merge, clean and upadte merged_1422_result dataset
```{r}
# Step 1: Aggregate FX_Rate by taking the mean and retain other columns
merged_1422_result <- merged_1422_result %>%
  group_by(Fiscal_Year, Country, HE.Provider, Level.Of.Study, Mode.Of..Study, 
           Academic.Year, Visa.Code, Visa.Fee...., Visa.NHS.Charge...., 
           Visa.After.Graduation, Visa.Work.Requirements.After.Graduation, 
           Salary.Requirements, Dependents.Allowed, Visa.Fee..Dependents, 
           Visa.NHS.Charge.Depen, HE.Provider.Term.1, HE.Provider.Term.2) %>%
  summarise(Enrollment = sum(Enrollment, na.rm = TRUE),
            FX_Rate = round(mean(FX_Rate, na.rm = TRUE), 2),  # Calculate mean FX_Rate and round to 2 decimal places
            .groups = 'drop') %>%
  distinct()  # Ensure no duplicates remain

# View the cleaned data
head(merged_1422_result)
```


#Merge 2 ( Combine the merged dataset with GDP and Inflation Rates)
```{r}
#adding the GDP/Inflation Dataset
gdp_inf_1422 <- read.csv("C:/Users/gochi/Desktop/CMM-500/Project Data/gdp_inf_semi_clean.csv" , stringsAsFactors = T)

#Reshape the GDP/Inflation dataset (pivot longer)
gdp_inf_long_1422 <- gdp_inf_1422 %>%
  pivot_longer(cols = starts_with("X"),  # Adjust for columns like X2014, X2015
               names_to = "Fiscal_Year", 
               values_to = "Value") %>%
  mutate(Year = sub("X", "", Fiscal_Year)) %>%  # Remove the "X" from the year
  pivot_wider(names_from = Growth.Indicator, values_from = Value)

  gdp_inf_long_1422$Fiscal_Year = NULL
  
  gdp_inf_long_1422 <- gdp_inf_long_1422 %>%
  rename(Fiscal_Year = Year)
#Reshape the GDP/Inflation dataset (pivot longer)
gdp_inf_long_1422 <- gdp_inf_1422 %>%
  pivot_longer(cols = starts_with("X"),  # Adjust for columns like X2014, X2015
               names_to = "Fiscal_Year", 
               values_to = "Value") %>%
  mutate(Fiscal_Year = sub("X", "", Fiscal_Year))  # Remove the "X" from the year


head(gdp_inf_long_1422)
```


```{r}

gdp_inf_long_1422$Fiscal_Year= as.factor(gdp_inf_long_1422$Fiscal_Year)
# Merge the datasets on Country and Year
merged_final_1422 <- merged_1422_result %>%
  left_join(gdp_inf_long_1422, by = c("Country", "Fiscal_Year"))

# Remove rows where Mode.Of..Study is "All"
merged_final_1422 <- merged_final_1422 %>%
  filter(Mode.Of..Study != "All")

#Drop the "All" level from the Mode.Of..Study factor
merged_final_1422 <- merged_final_1422 %>%
  mutate(Mode.Of..Study = droplevels(Mode.Of..Study))

# View the cleaned data
summary(merged_final_1422)
```


#Investigating Duplicates
```{r}
# Step 1: Check for duplicates in the enrollment dataset
enrollment_duplicates <- merged_1422_result %>%
  group_by(Country, Fiscal_Year) %>%
  filter(n() > 1)

# View duplicates in the enrollment dataset
print(enrollment_duplicates)

# Step 2: Check for duplicates in the GDP/Inflation dataset
gdp_inflation_duplicates <- gdp_inf_long_1422 %>%
  group_by(Country,Fiscal_Year) %>%
  filter(n() > 1)

# View duplicates in the GDP/Inflation dataset
print(gdp_inflation_duplicates)
#Save or view the result
write.csv(merged_final_1422, "C:/Users/gochi/Desktop/CMM-500/Project Data/merge_final_1422.csv", row.names = FALSE)
```

##still processing ----
```{r}
#Aggregate Value by taking the mean and retain other columns
merged_final_1422 <- merged_final_1422 %>%
  group_by(Fiscal_Year, Country, HE.Provider, Level.Of.Study, Mode.Of..Study, 
           Academic.Year, Visa.Code, Visa.Fee...., Visa.NHS.Charge...., 
           Visa.After.Graduation, Visa.Work.Requirements.After.Graduation, 
           Salary.Requirements, Dependents.Allowed, Visa.Fee..Dependents, 
           Visa.NHS.Charge.Depen, HE.Provider.Term.1, HE.Provider.Term.2, 
           Growth.Indicator) %>%  # Include Growth.Indicator in the group_by
  summarise(Enrollment = sum(Enrollment, na.rm = TRUE),
            Value = round(mean(Value, na.rm = TRUE), 2),  # Calculate mean Value and round to 2 decimal places
            .groups = 'drop') %>%
  distinct()  # Ensure no duplicates remain

# View the cleaned data
head(merged_final_1422)
```

```{r}
merged_final_1422$HE.Provider.Term.1 = NULL
merged_final_1422$HE.Provider.Term.2 = NULL
summary(merged_final_1422)
```

#---final merge with tuition

##Add tuition Fees Data
```{r}
#Load the dataset
tuition_1422_cleaned <- read.csv("C:/Users/gochi/Desktop/CMM-500/Project Data/tuition_long1422.csv" , stringsAsFactors = T)
tuition_1422_cleaned$Fiscal_Year = as.factor(tuition_1422_cleaned$Fiscal_Year)

# Merge the datasets on common keys ( Fiscal_Year, and HE.Provider)
final_merged_data_tuition <- merged_final_1422 %>%
  left_join(tuition_1422_cleaned, by = c("HE.Provider", "Fiscal_Year"))
```


```{r}
final_merged_data_tuition$Country = as.factor(final_merged_data_tuition$Country)
final_merged_data_tuition$HE.Provider = as.factor(final_merged_data_tuition$HE.Provider)
summary(final_merged_data_tuition)
#write.csv(final_merged_data_tuition,"C:/Users/gochi/Desktop/CMM-500/Project Data/final_merged_data_tuition.csv" ,row.names = FALSE)

```
##University of Aberdeen has duplicates
```{r}
# Standardise HE.Provider names by trimming spaces and fixing known discrepancies
final_merged_data_tuition <- final_merged_data_tuition %>%
  mutate(HE.Provider = str_trim(HE.Provider))  # Trim leading and trailing spaces

# Step 2: Correct specific cases where "University of Aberdeen" might be misspelled or have extra spaces
final_merged_data_tuition <- final_merged_data_tuition %>%
  mutate(HE.Provider = case_when(
    str_detect(HE.Provider, regex("The\\s*University\\s*of\\s*Aberdeen", ignore_case = TRUE)) ~ "The University of Aberdeen",
    TRUE ~ HE.Provider
  ))

# Step 3: Remove duplicates for "University of Aberdeen"
aberdeen_data_cleaned <- final_merged_data_tuition %>%
  filter(HE.Provider == "The University of Aberdeen") %>%
  distinct(Fiscal_Year, Country, HE.Provider, Level.Of.Study, Mode.Of..Study, 
           Academic.Year, Growth.Indicator, Value, .keep_all = TRUE)

# Step 4: Combine the cleaned "University of Aberdeen" data with the rest of the dataset
  final_merged_data_tuition <- final_merged_data_tuition %>%
  filter(HE.Provider != "The University of Aberdeen") %>%  # Keep all other universities
  bind_rows(aberdeen_data_cleaned)  # Add the cleaned "University of Aberdeen" data back

  final_merged_data_tuition$HE.Provider = as.factor(final_merged_data_tuition$HE.Provider)
# View the cleaned data
head(final_merged_data_tuition)
```

```{r}
summary(final_merged_data_tuition$HE.Provider)
summary(final_merged_data_tuition)
```


```{r}
final_merged_data_tuition_clean = final_merged_data_tuition
final_merged_data_tuition_clean$HE.Provider = as.factor(final_merged_data_tuition_clean$HE.Provider)
final_merged_data_tuition_clean$Visa.Fee.... = as.factor(final_merged_data_tuition_clean$Visa.Fee....)
final_merged_data_tuition_clean$Visa.NHS.Charge....= as.factor(final_merged_data_tuition_clean$Visa.NHS.Charge....)
final_merged_data_tuition_clean$Visa.Fee..Dependents =as.factor(final_merged_data_tuition_clean$Visa.Fee..Dependents)
final_merged_data_tuition_clean$Visa.NHS.Charge.Depen= as.factor(final_merged_data_tuition_clean$Visa.NHS.Charge.Depen)
summary(final_merged_data_tuition_clean)
#write.csv(final_merged_data_tuition_clean,"C:/Users/gochi/Desktop/CMM-500/Project Data/final_merged_data_tuition_clean.csv" ,row.names = FALSE)
```
Noticed that FX rate is missing! need to add it 
-----------------------
```{r}
# Rename the Year column to Fiscal_Year
fx_long_1422 <- fx_long_1422 %>%
  rename(Fiscal_Year = Year)
fx_long_1422$Fiscal_Year = as.factor(fx_long_1422$Fiscal_Year)
```


```{r}
# Aggregate FX_Rate by taking the mean for each Fiscal_Year and Country combination
fx_long_1422_cleaned <- fx_long_1422 %>%
  group_by(Fiscal_Year, Country) %>%
  summarise(FX_Rate = mean(FX_Rate, na.rm = TRUE))

# View the cleaned FX data
head(fx_long_1422_cleaned)
```


```{r}
# Merge the datasets by Fiscal_Year and Country
merged_data_check <- left_join(final_merged_data_tuition_clean, fx_long_1422_cleaned, by = c("Fiscal_Year", "Country"))
merged_data_check$Country = as.factor(merged_data_check$Country)
summary(merged_data_check)
```


renaming datasets
```{r}
final_combined_1422 = merged_data_check
#write.csv(final_combined_1422,"C:/Users/gochi/Desktop/CMM-500/Project Data/final_combined_1422.csv" ,row.names = FALSE)
```



```{r}
#convert some factors to numerical values
final_combined_1422$Visa.Fee.... =as.factor(final_combined_1422$Visa.Fee....)
final_combined_1422$Visa.NHS.Charge....=as.factor(final_combined_1422$Visa.NHS.Charge....)
final_combined_1422$Visa.Fee..Dependents= as.factor(final_combined_1422$Visa.Fee..Dependents)
final_combined_1422$Visa.NHS.Charge.Depen= as.factor(final_combined_1422$Visa.NHS.Charge.Depen)
```

##=================================================================================DATA EXPLORATION ANALYSIS====================================================================================

```{r}
summary(final_combined_1422)
#summary(final_combined_1422$Fiscal_Year)
```


Summary
Time Frame:
The data spans from the fiscal year 2014 to 2021 and academic year 2014/2015 - 2021/2022.

Countries:
The dataset includes students from a variety of countries (15), with the most frequent being China, Gambia, Ghana, India, Japan, and Kenya.

Higher Education Providers:
The dataset includes multiple universities in Scotland, United Kindgom

The majority of the records are grouped under the "Other" category, which includes various other institutions.
Level and Mode of Study:
The dataset primarily focuses on postgraduate students.
The mode of study is split between full-time and part-time, with a fairly even distribution between the two.

Visa Information:
Visa Codes:
There are four types of visa codes listed:
Student Visa
Tier 4 (General)
Tier 4 + Graduate Route
Tier 4 Pilot

Visa Fees:
Visa fees range between £310 and £348, depending on the type and year.
NHS Surcharge:
The NHS surcharge has three levels: £150, £300, and £470.

Post-Graduation Work Requirements:
Different work requirements after graduation are noted:
Some students can stay under company-dependent conditions.
Graduate Route allows for a 2-year post-study work period.
Some have no work requirements.
Salary Requirements:
A significant portion of records requires a minimum salary of £20,800, while others do not have any specified salary requirement.

Growth Indicators:
The dataset tracks economic indicators such as GDP and Inflation, with an equal distribution among the records.

Enrollment:
Enrollment numbers vary significantly, with a minimum of 0 and a maximum of 97,140. The mean enrollment is around 314.7, indicating that some institutions or programs have large enrollments, while others have very few students.

Value(GDP and Inflation):
The "Value" column ranges from a minimum of -58.32 to a maximum of 153.49. The negative values might indicate a financial loss or deficit, while positive values likely indicate some form of gain or contribution.

Tuition Fees:
Tuition fees for different programs and institutions range from £12,000 to £23,250, with an average of around £16,608.

Key Observations:
Visa Policy Trends: Over the years, visa policies have shifted, with more stringent requirements and varying fees.

Enrollment Variability: Enrollment numbers vary widely, suggesting some programs or universities are much more popular than others.

Tuition Fee Distribution: Tuition fees have a broad range, possibly reflecting differences in the types of programs or the prestige of the institutions.


```{r}
summary(final_combined_1422$HE.Provider)
```

------------------------------------Distributions and Visualisations----------------------------------------------------------
Visaulising Numerical Variables
1. Enrolment
2. GDP Values
3. Inflation Values
5. Tuition Fees
6. Visa Fees


```{r,echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=4}
ggplot(data = final_combined_1422, aes(x = Enrollment)) + geom_histogram(bins = 8) #enrolment numbers
ggplot(data = final_combined_1422, aes(x = Enrollment)) + geom_density(col = 'blue', lwd = 0.5) #enrolment numbers
ggplot(data = final_combined_1422, aes(x = FX_Rate)) + geom_density(col = 'red', lwd = 0.5) #fx_rate
ggplot(data = final_combined_1422, aes(x = Value)) + geom_density(col = 'blue', lwd = 0.5)
ggplot(data = final_combined_1422, aes(x = Value)) + geom_density(col = 'blue', lwd = 0.5) + facet_wrap(~ Growth.Indicator)# see both GDP and Inflation + Negative distribution
ggplot(data = final_combined_1422, aes(x = "", y = Value)) + geom_violin() # to see the negative values
ggplot(data = final_combined_1422, aes(x = Growth.Indicator, y = Value, fill = Growth.Indicator)) + geom_bar(stat = "identity") + labs(x = "Growth Indicator Rate", y = "Value %")
```

Distribution of Categorical Variables
---------------------------
Mode of Study 
Visa Code
HE Provider 
Growth Indicator
Country

```{r,=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
ggplot(final_combined_1422, aes(x = Visa.Code , fill= Visa.Code)) + geom_bar() + labs (" Visa Codes Class Distribution")
ggplot(final_combined_1422, aes(x = Mode.Of..Study , fill= Mode.Of..Study)) + geom_bar() + labs("Modes of Study Class Distribution")
ggplot(final_combined_1422, aes(x = HE.Provider,y = "")) + geom_bar( stat = "identity") + coord_flip() + labs(title = "HE Provider Distribution")
ggplot(final_combined_1422, aes(x = Growth.Indicator,fill = Growth.Indicator)) + geom_bar() + labs ( title = "Class Distribution of Growth Indicators")
ggplot(final_combined_1422, aes(x = Country)) + geom_bar()
```

Distribution in Pairs
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
custom_palette <- colorRampPalette(brewer.pal(12, "Set3"))(18)
ggplot(data = final_combined_1422, aes(x = Country, y = FX_Rate, fill = Country)) + geom_bar(stat = "identity") + labs(x = "Country", y = "FX Rate Value")
ggplot(data = final_combined_1422, aes(x = Country, y = Value, fill = Country)) + geom_bar(stat = "identity") + labs(x = "Country", y = "Growth Indicator Rate")
#ggplot(data = final_combined_1422, aes(x = Enrollment, y = Country, fill = HE.Provider)) + geom_bar(stat = "identity") + labs(x = "Enrolment numbers", y = "Country of Origin") + scale_fill_manual(values = custom_palette)
ggplot(data = final_combined_1422, aes(x = Country, y = Enrollment, fill = Country)) + geom_bar(stat = "identity") + labs(x = "Country", y = "Enrollment")
ggplot(data = final_combined_1422, aes(x = HE.Provider, y = Enrollment, fill = Country)) + geom_bar(stat = "identity") + labs(x = "HE.Provider", y = "Enrollment") + coord_flip() 
```


##---Treating Outliers-----------------------------

```{r}
# Identify numeric and integer columns
numeric_columns <- sapply(final_combined_1422, is.numeric)
int_columns <- sapply(final_combined_1422, is.integer)

# Combine both into one logical vector
num_int_columns <- numeric_columns | int_columns

# Extract just the numeric and integer columns
numeric_data <- final_combined_1422[, num_int_columns]

#summary(numeric_data)

```

```{r}
# Reshape data for plotting
numeric_data_melted <- melt(numeric_data)

# Create boxplots
ggplot(numeric_data_melted, aes(x = variable, y = value)) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_y_continuous(trans='log10') +  # Apply log scale on y-axis
  labs(title = "Boxplots to Identify Outliers", x = "Variable", y = "Log(Value + 1)")

```

Detecting the outliers
```{r}
# Function to detect outliers
detect_outliers <- function(x) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr <- q3 - q1
  lower_bound <- q1 - 1.5 * iqr
  upper_bound <- q3 + 1.5 * iqr
  x[x < lower_bound | x > upper_bound]
}

# Apply the function to each numeric/int column
outliers <- lapply(numeric_data, detect_outliers)

# View outliers
outliers
summary(outliers)
summary(outliers$Enrollment)
summary(outliers$FX_Rate)
summary(outliers$Value)
summary(outliers$Tuition_Fees)
```


#removing outliers

```{r}
 # Define a function to remove outliers and return the clipped values
  remove_outliers <- function(df, column) {
  Q1 <- quantile(df[[column]], 0.25, na.rm = TRUE)
  Q3 <- quantile(df[[column]], 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  
  # Define lower and upper bounds
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  
  # Clip values outside the bounds by setting them to NA
  df[[column]] <- ifelse(df[[column]] < lower_bound | df[[column]] > upper_bound, NA, df[[column]])
  
  return(df[[column]])  # Return the modified column
}

# Apply the clipping function to FX_Rate, Value, Enrollment, and Tuition_Fees
final_combined_1422_no_outlier <- final_combined_1422
final_combined_1422_no_outlier$FX_Rate <- remove_outliers(final_combined_1422_no_outlier, "FX_Rate")
final_combined_1422_no_outlier$Value <- remove_outliers(final_combined_1422_no_outlier, "Value")
final_combined_1422_no_outlier$Enrollment <- remove_outliers(final_combined_1422_no_outlier, "Enrollment")
#final_combined_1422_no_outlier$Tuition_Fees <- remove_outliers(final_combined_1422_no_outlier, "Tuition_Fees")

# Optionally, remove rows with NAs introduced by outlier removal
final_combined_1422_no_outlier <- final_combined_1422_no_outlier %>%
  drop_na(FX_Rate, Value, Enrollment)

# View the cleaned data summary
summary(final_combined_1422_no_outlier)


```
Visualisation after removing outliers
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=4}
ggplot(data = final_combined_1422_no_outlier, aes(x = Enrollment)) + geom_histogram(bins = 8) #enrolment numbers
ggplot(data = final_combined_1422_no_outlier, aes(x = Enrollment)) + geom_density(col = 'blue', lwd = 0.5) #enrolment numbers
ggplot(data = final_combined_1422_no_outlier, aes(x = FX_Rate)) + geom_density(col = 'red', lwd = 0.5) #fx_rate
ggplot(data = final_combined_1422_no_outlier, aes(x = Value)) + geom_density(col = 'blue', lwd = 0.5)
ggplot(data = final_combined_1422_no_outlier, aes(x = Value)) + geom_density(col = 'blue', lwd = 0.5) + facet_wrap(~ Growth.Indicator)# see both GDP and Inflation + Negative distribution
ggplot(data = final_combined_1422_no_outlier, aes(x = "", y = Value)) + geom_violin() # to see the negative values
ggplot(data = final_combined_1422_no_outlier, aes(x = Growth.Indicator, y = Value, fill = Growth.Indicator)) + geom_bar(stat = "identity") + labs(x = "Growth Indicator Rate", y = "Value %")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
custom_palette <- colorRampPalette(brewer.pal(12, "Set3"))(18)
ggplot(data = final_combined_1422_no_outlier, aes(x = Country, y = FX_Rate, fill = Country)) + geom_bar(stat = "identity") + labs(x = "Country", y = "FX Rate Value")
ggplot(data = final_combined_1422_no_outlier, aes(x = Country, y = Value, fill = Country)) + geom_bar(stat = "identity") + labs(x = "Country", y = "Growth Indicator Rate")
#ggplot(data = final_combined_1422, aes(x = Enrollment, y = Country, fill = HE.Provider)) + geom_bar(stat = "identity") + labs(x = "Enrolment numbers", y = "Country of Origin") + scale_fill_manual(values = custom_palette)
ggplot(data = final_combined_1422_no_outlier, aes(x = Country, y = Enrollment, fill = Country)) + geom_bar(stat = "identity") + labs(x = "Country", y = "Enrollment")
ggplot(data = final_combined_1422_no_outlier, aes(x = HE.Provider, y = Enrollment, fill = Country)) + geom_bar(stat = "identity") + labs(x = "HE.Provider", y = "Enrollment") + coord_flip() 
```

Clipping the outliers to compare with removed outliers and normal dataset
```{r}
# Define a function for percentile-based clipping
clip_outliers <- function(x, lower_percentile = 0.05, upper_percentile = 0.95) {
  lower_bound <- quantile(x, lower_percentile, na.rm = TRUE)
  upper_bound <- quantile(x, upper_percentile, na.rm = TRUE)
  
  x[x < lower_bound] <- lower_bound
  x[x > upper_bound] <- upper_bound
  
  return(x)
}

# Apply the clipping function to FX_Rate, Value, and Enrollment
final_combined_1422_clipped <- final_combined_1422
final_combined_1422_clipped$FX_Rate <- clip_outliers(final_combined_1422$FX_Rate)
final_combined_1422_clipped$Value <- clip_outliers(final_combined_1422$Value)
final_combined_1422_clipped$Enrollment <- clip_outliers(final_combined_1422$Enrollment)
```

See Visualisations of the dataset after clipping at 95% 
```{r}
ggplot(data = final_combined_1422_clipped, aes(x = Enrollment)) + geom_histogram(bins = 8) #enrolment numbers
ggplot(data = final_combined_1422_clipped, aes(x = Enrollment)) + geom_density(col = 'blue', lwd = 0.5) #enrolment numbers
ggplot(data = final_combined_1422_clipped, aes(x = FX_Rate)) + geom_density(col = 'red', lwd = 0.5) #fx_rate
ggplot(data = final_combined_1422_clipped, aes(x = Value)) + geom_density(col = 'blue', lwd = 0.5)
ggplot(data = final_combined_1422_clipped, aes(x = Value)) + geom_density(col = 'blue', lwd = 0.5) + facet_wrap(~ Growth.Indicator)# see both GDP and Inflation + Negative distribution
ggplot(data = final_combined_1422_clipped, aes(x = "GDP/Inflation", y = Value)) + geom_violin() # to see the negative values
ggplot(data = final_combined_1422_clipped, aes(x = Growth.Indicator, y = Value, fill = Growth.Indicator)) + geom_bar(stat = "identity") + labs(x = "Growth Indicator Rate", y = "Value")
```
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
custom_palette <- colorRampPalette(brewer.pal(12, "Set3"))(18)
ggplot(data = final_combined_1422_clipped, aes(x = Country, y = FX_Rate, fill = Country)) + geom_bar(stat = "identity") + labs(x = "Country", y = "FX Rate Value")
ggplot(data = final_combined_1422_clipped, aes(x = Country, y = Value, fill = Country)) + geom_bar(stat = "identity") + labs(x = "Country", y = "Growth Indicator Rate")
#ggplot(data = final_combined_1422, aes(x = Enrollment, y = Country, fill = HE.Provider)) + geom_bar(stat = "identity") + labs(x = "Enrolment numbers", y = "Country of Origin") + scale_fill_manual(values = custom_palette)
ggplot(data = final_combined_1422_clipped, aes(x = Country, y = Enrollment, fill = Country)) + geom_bar(stat = "identity") + labs(x = "Country", y = "Enrollment")
ggplot(data = final_combined_1422_clipped, aes(x = HE.Provider, y = Enrollment, fill = Country)) + geom_bar(stat = "identity") + labs(x = "HE.Provider", y = "Enrollment") + coord_flip() 
```


Summary of Numeric variables after outliers have been clipped vs removed vs original
```{r}
# Summarize FX_Rate
cat("Summary of FX_Rate for Original, Clipped, and No Outliers Datasets:\n")
summary(final_combined_1422$FX_Rate)
cat("\n---\n")
summary(final_combined_1422_clipped$FX_Rate)
cat("\n---\n")
summary(final_combined_1422_no_outlier$FX_Rate)

cat("\n===================================\n")

# Summarize Value
cat("Summary of Value for Original, Clipped, and No Outliers Datasets:\n")
summary(final_combined_1422$Value)
cat("\n---\n")
summary(final_combined_1422_clipped$Value)
cat("\n---\n")
summary(final_combined_1422_no_outlier$Value)

cat("\n===================================\n")

# Summarize Enrollment
cat("Summary of Enrollment for Original, Clipped, and No Outliers Datasets:\n")
summary(final_combined_1422$Enrollment)
cat("\n---\n")
summary(final_combined_1422_clipped$Enrollment)
cat("\n---\n")
summary(final_combined_1422_no_outlier$Enrollment)
```
```{r}
#visualise
# Create a new column to identify the dataset type
final_combined_1422$Dataset <- "Original"
final_combined_1422_clipped$Dataset <- "Clipped"
final_combined_1422_no_outlier$Dataset <- "No Outliers"

# Combine the three datasets into one
combined_data <- bind_rows(final_combined_1422, final_combined_1422_clipped, final_combined_1422_no_outlier)

# Create the boxplot for Enrollment with Log Scale, showing each dataset side by side
ggplot(combined_data, aes(x = Dataset, y = log(FX_Rate + 1), fill = Dataset)) + 
  geom_boxplot() +
  labs(title = "FX_Rate: Original vs Clipped vs No Outliers [Log Scale]", y = "Log(FXRate + 1)", x = "Dataset") +
  theme_minimal() +
  scale_fill_manual(values = c("Original" = "blue", "Clipped" = "red", "No Outliers" = "green"))

#----------------------

# Create the boxplot for Enrollment with Log Scale, showing each dataset side by side
ggplot(combined_data, aes(x = Dataset, y = log(Enrollment + 1), fill = Dataset)) + 
  geom_boxplot() +
  labs(title = "Enrollment: Original vs Clipped vs No Outliers [Log Scale]", y = "Log(Enrollment + 1)", x = "Dataset") +
  theme_minimal() +
  scale_fill_manual(values = c("Original" = "blue", "Clipped" = "red", "No Outliers" = "green"))

#-----------------------------
# Create the boxplot for Value, showing each dataset side by side
ggplot(combined_data, aes(x = Dataset, y = log(Value + 1), fill = Dataset)) + 
  geom_boxplot() +
  labs(title = "Growth Indicators: Original vs Clipped vs No Outliers [Log Scale]", y = "Log(Value + 1)", x = "Dataset") +
  theme_minimal() +
  scale_fill_manual(values = c("Original" = "blue", "Clipped" = "red", "No Outliers" = "green"))

```

## ---------------FINANCIAL TRENDS--
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
final_combined_1422_no_outlier$Fiscal_Year= as.numeric(final_combined_1422_no_outlier$Fiscal_Year)
final_combined_1422_clipped$Fiscal_Year= as.numeric(final_combined_1422_clipped$Fiscal_Year)
# Plot FX Rate over time by Country
ggplot(final_combined_1422_no_outlier, aes(x = Fiscal_Year, y = FX_Rate, color = Country)) +
  geom_line() +
  labs(title = "FX Rate Over Time by Country - no outlier", x = "Fiscal Year", y = "FX Rate") +
  facet_wrap(~ Country, scales = "free_y", ncol = 3) +  # Separate plots by country with free y-axis scaling
  theme_minimal()

# Plot GDP/Inflation Value over time by Country
ggplot(final_combined_1422_no_outlier, aes(x = Fiscal_Year, y = Value, color = Country)) +
  geom_line() +
  labs(title = "GDP/Inflation Value Over Time by Country -no outlier", x = "Fiscal Year", y = "GDP/Inflation Value") +
  facet_wrap(~ Country, scales = "free_y", ncol = 3) +  # Separate plots by country with free y-axis scaling
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}

# Plot FX Rate over time by Country
ggplot(final_combined_1422_clipped, aes(x = Fiscal_Year, y = FX_Rate, color = Country)) +
  geom_line() +
  labs(title = "FX Rate Over Time by Country - Clipped ", x = "Fiscal Year", y = "FX Rate") +
  facet_wrap(~ Country, scales = "free_y", ncol = 3) +  # Separate plots by country with free y-axis scaling
  theme_minimal()

# Plot GDP/Inflation Value over time by Country
ggplot(final_combined_1422_clipped, aes(x = Fiscal_Year, y = Value, color = Country)) +
  geom_line() +
  labs(title = "GDP/Inflation Value Over Time by Country -Clipped", x = "Fiscal Year", y = "GDP/Inflation Value") +
  facet_wrap(~ Country, scales = "free_y", ncol = 3) +  # Separate plots by country with free y-axis scaling
  theme_minimal()
```

## ENROLMENT TRENDS-----

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
#-------------------------------------------CLIPPED-----------------------------------------------
# Convert Fiscal_Year to a numeric format if it isn't already
final_combined_1422_clipped$Fiscal_Year <- as.numeric(as.character(final_combined_1422_clipped$Fiscal_Year))
#-----
final_combined_1422_clipped <- final_combined_1422_clipped %>%
  group_by(Country) %>%
  mutate(Rolling_Average = zoo::rollmean(Enrollment, k = 3, fill = NA))  # k=3 for a 3-period moving average

ggplot(final_combined_1422_clipped, aes(x = Fiscal_Year, y = Rolling_Average, color = Country)) +
  geom_line(size = 1) +
  labs(title = "3-Year Rolling Average of Enrollment by Country - Clipped",
       x = "Fiscal Year",
       y = "Rolling Average Enrollment") +
  theme_minimal()
#-----
ggplot(final_combined_1422_clipped, aes(x = Fiscal_Year, y = Enrollment, color = Country)) +
  geom_line(size = 1) +
  facet_wrap(~ Country, scales = "free_y") +  # Creates separate panels for each country
  labs(title = "Student Enrollment Over Time by Country - Clipped",
       x = "Fiscal Year",
       y = "Enrollment") +
  theme_minimal()
```


```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=9}
#-------------------------------------------NO OUTLIER-----------------------------------------------
# Convert Fiscal_Year to a numeric format if it isn't already
final_combined_1422_no_outlier$Fiscal_Year <- as.numeric(as.character(final_combined_1422_no_outlier$Fiscal_Year))
#-----
final_combined_1422_no_outlier <- final_combined_1422_no_outlier %>%
  group_by(Country) %>%
  mutate(Rolling_Average = zoo::rollmean(Enrollment, k = 3, fill = NA))  # k=3 for a 3-period moving average

ggplot(final_combined_1422_clipped, aes(x = Fiscal_Year, y = Rolling_Average, color = Country)) +
  geom_line(size = 1) +
  labs(title = "3-Year Rolling Average of Enrollment by Country - No- Outlier",
       x = "Fiscal Year",
       y = "Rolling Average Enrollment") +
  theme_minimal()
#-----
ggplot(final_combined_1422_no_outlier, aes(x = Fiscal_Year, y = Enrollment, color = Country)) +
  geom_line(size = 1) +
  facet_wrap(~ Country, scales = "free_y") +  # Creates separate panels for each country
  labs(title = "Student Enrollment Over Time by Country - No Outlier",
       x = "Fiscal Year",
       y = "Enrollment") +
  theme_minimal()
```

##-----TUITION FEES TRENDS--------------------

```{r}

##-------------------CLIPPED---------------------
##Examine the distribution of tuition fees to understand their range and potential impact on enrollment
# Visualise the distribution of Tuition Fees
ggplot(final_combined_1422_clipped, aes(x = Tuition_Av)) + 
  geom_histogram(binwidth = 500, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Tuition Fees- Clipped", x = "Tuition Fees", y = "Frequency")

# Density plot
ggplot(final_combined_1422_clipped, aes(x = Tuition_Av)) + 
  geom_density(fill = "lightgreen") +
  labs(title = "Density Plot of Tuition Fees - Clipped", x = "Tuition Fees", y = "Density")

# Correlation with Enrollment
tuition_enrollment_corr <- cor(final_combined_1422_clipped$Tuition_Av, final_combined_1422_clipped$Enrollment, method = "spearman")
print(tuition_enrollment_corr)

# Scatter plot to visualise relationship
ggplot(final_combined_1422_clipped, aes(x = Tuition_Av, y = Enrollment)) +
  geom_point(aes(color = Country), alpha = 0.6) +
  labs(title = "Tuition Fees vs Enrollment -Clipped", x = "Tuition Fees", y = "Enrollment") +
  theme_minimal()
```

```{r}
#----- NO OUTLIER----------------
##Examine the distribution of tuition fees to understand their range and potential impact on enrollment
# Visualise the distribution of Tuition Fees
ggplot(final_combined_1422_no_outlier, aes(x = Tuition_Av)) + 
  geom_histogram(binwidth = 500, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Tuition Fees- No Outlier", x = "Tuition Fees", y = "Frequency")

# Density plot
ggplot(final_combined_1422_no_outlier, aes(x = Tuition_Av)) + 
  geom_density(fill = "lightgreen") +
  labs(title = "Density Plot of Tuition Fees - No Outlier", x = "Tuition Fees", y = "Density")

# Correlation with Enrollment
tuition_enrollment_corr2 <- cor(final_combined_1422_no_outlier$Tuition_Av, final_combined_1422_no_outlier$Enrollment, method = "spearman")
print(tuition_enrollment_corr2)

# Scatter plot to visualise relationship
 ggplot(final_combined_1422_no_outlier, aes(x = Tuition_Av, y = Enrollment)) +
  geom_point(aes(color = Country), alpha = 0.6) +
  labs(title = "Tuition Fees vs Enrollment - No Outlier", x = "Tuition Fees", y = "Enrollment") +
  theme_minimal()
```
##---- VISA POLICY CHHANGES OVER TIME-------------------------------------


```{r}
#------ No Outliers----------------------
# Count the frequency of Visa Policy Changes by Year
visa_policy_changes <- final_combined_1422_no_outlier %>%
  group_by(Fiscal_Year, Visa.Code) %>%
  summarise(Change_Frequency = n())

# Plot the frequency of visa policy changes over time
ggplot(visa_policy_changes, aes(x = Fiscal_Year, y = Change_Frequency, fill = Visa.Code)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Frequency of Visa Policy Changes by Year - no outliers", x = "Year", y = "Frequency") +
  theme_minimal()

# Enrollment trends with Visa Policy Changes
ggplot(final_combined_1422_no_outlier, aes(x = Fiscal_Year, y = Enrollment, group = Visa.Code, color = Visa.Code)) +
  geom_line() +
  geom_point() +
  labs(title = "Enrollment Trends with Visa Policy Changes - no outliers", x = "Year", y = "Enrollment") +
  theme_minimal()
```
```{r}
#------ Clipped----------------------
# Count the frequency of Visa Policy Changes by Year
visa_policy_changes <- final_combined_1422_clipped %>%
  group_by(Fiscal_Year, Visa.Code) %>%
  summarise(Change_Frequency = n())

# Plot the frequency of visa policy changes over time
ggplot(visa_policy_changes, aes(x = Fiscal_Year, y = Change_Frequency, fill = Visa.Code)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Frequency of Visa Policy Changes by Year - clipped", x = "Year", y = "Frequency") +
  theme_minimal()

# Enrollment trends with Visa Policy Changes
ggplot(final_combined_1422_clipped, aes(x = Fiscal_Year, y = Enrollment, group = Visa.Code, color = Visa.Code)) +
  geom_line() +
  geom_point() +
  labs(title = "Enrollment Trends with Visa Policy Changes - clipped", x = "Year", y = "Enrollment") +
  theme_minimal()
```

##--- VISA POLICIES AGAINST ENROLMENT
```{r}
# Assuming your dataset is called 'final_combined_1422_clipped'
# and has columns 'Year', 'Enrollment', 'Visa.Code'

# Create a basic line plot of enrollment over time by visa type
ggplot(final_combined_1422_clipped, aes(x = Fiscal_Year, y = Enrollment, color = Visa.Code, group = Visa.Code)) +
  geom_line(size = 1) +  # Line plot for enrollment over time
  geom_point(size = 2) +  # Adding points for clarity
  scale_y_continuous(labels = scales::comma) +  # Format y-axis labels as comma separated
  labs(title = "Enrollment Figures Against Visa Changes Over Time",
       x = "Year",
       y = "Enrollment",
       color = "Visa Code") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


#Summary statistics (e.g., mean, median, standard deviation) for enrollment across different countries, universities and modes of study.

```{r}

  # Calculate summary statistics for enrollment by Country
  enrollment_by_country <- final_combined_1422_clipped %>%
  group_by(Country) %>%
  summarise(
    Mean_Enrollment = mean(Enrollment, na.rm = TRUE),
    Median_Enrollment = median(Enrollment, na.rm = TRUE),
    SD_Enrollment = sd(Enrollment, na.rm = TRUE),
    Min_Enrollment = min(Enrollment, na.rm = TRUE),
    Max_Enrollment = max(Enrollment, na.rm = TRUE),
    Total_Enrollment = sum(Enrollment, na.rm = TRUE)
  )

# Calculate summary statistics for enrollment by University
enrollment_by_university <- final_combined_1422_clipped %>%
  group_by(HE.Provider) %>%
  summarise(
    Mean_Enrollment = mean(Enrollment, na.rm = TRUE),
    Median_Enrollment = median(Enrollment, na.rm = TRUE),
    SD_Enrollment = sd(Enrollment, na.rm = TRUE),
    Min_Enrollment = min(Enrollment, na.rm = TRUE),
    Max_Enrollment = max(Enrollment, na.rm = TRUE),
    Total_Enrollment = sum(Enrollment, na.rm = TRUE)
  )

# Calculate summary statistics for enrollment by Mode of Study
enrollment_by_mode <- final_combined_1422_clipped %>%
  group_by(Mode.Of..Study) %>%
  summarise(
    Mean_Enrollment = mean(Enrollment, na.rm = TRUE),
    Median_Enrollment = median(Enrollment, na.rm = TRUE),
    SD_Enrollment = sd(Enrollment, na.rm = TRUE),
    Min_Enrollment = min(Enrollment, na.rm = TRUE),
    Max_Enrollment = max(Enrollment, na.rm = TRUE),
    Total_Enrollment = sum(Enrollment, na.rm = TRUE)
  )

# Display the results
enrollment_by_country
enrollment_by_university
enrollment_by_mode
```

---##Currency and Economic Relationships:
```{r}
# The plot can be enhanced by color-coding the points based on values of a 3rd variable:
  ggplot(data = final_combined_1422_clipped, aes(x = FX_Rate, y = Enrollment, color = Tuition_Av)) + 
  geom_point() +
  scale_color_gradient(low = "blue", high = "cyan") +  # Color scale from blue to red based on Tuition Fees
  labs(title = "Enrollment vs FX_Rate with Tuition Fees - Clipped",
       x = "FX_Rate",
       y = "Enrollment") +
  theme_minimal()

# The plot can be enhanced by color-coding the points based on values of a 3rd variable:
  ggplot(data = final_combined_1422_no_outlier, aes(x = FX_Rate, y = Enrollment, color = Tuition_Av)) + 
  geom_point() +
  scale_color_gradient(low = "blue", high = "cyan") +  # Color scale from blue to red based on Tuition Fees
  labs(title = "Enrollment vs FX_Rate with Tuition Fees - No Outlier",
       x = "FX_Rate",
       y = "Enrollment") +
  theme_minimal()
  
  #adding other numerical variables
  ggplot(data = final_combined_1422_clipped, aes(x = FX_Rate, y = Enrollment, color = Value)) + 
  geom_point() +
  scale_color_gradient(low = "blue", high = "yellow") +  # Color scale from blue to red based on Tuition Fees
  labs(title = "Enrollment vs FX_Rate with Growth Indicators (GDP/Inflation Values) - Clipped",
       x = "FX_Rate",
       y = "Enrollment") +
  theme_minimal()
 
  ggplot(data = final_combined_1422_no_outlier, aes(x = FX_Rate, y = Enrollment, color = Value)) + 
  geom_point() +
  scale_color_gradient(low = "blue", high = "yellow") +  # Color scale from blue to red based on Tuition Fees
  labs(title = "Enrollment vs FX_Rate with Growth Indicators (GDP/Inflation Values) - No Outlier",
       x = "FX_Rate",
       y = "Enrollment") +
  theme_minimal()
 
```


Summary --

1. Enrollment vs FX_Rate with Tuition Fees - Clipped
Observation:
The majority of enrollment is concentrated on the left side of the plot where FX_Rates are relatively low (close to zero).
As FX_Rates increase, the number of enrollments decreases significantly.
Tuition fees also vary widely within this range, with a noticeable spread but concentrated more towards lower FX_Rates.
This suggests that higher FX_Rates may correlate with lower enrollments and potentially higher tuition fees.

2. Enrollment vs FX_Rate with Tuition Fees - No Outlier
Observation:
After removing outliers, the spread of data is much more condensed. The number of enrollments is still highest at low FX_Rates.
The plot reveals that enrollments are more consistently distributed across FX_Rates, though the concentration is still higher at lower FX_Rates.
This pattern reinforces the earlier observation that as FX_Rates increase, enrollments tend to decrease, though the trend is less pronounced without outliers.

3. Enrollment vs FX_Rate with Growth Indicators (GDP/Inflation Values) - Clipped
Observation:
The color gradient (representing growth indicators like GDP and Inflation) shows varying economic conditions across different FX_Rates and enrollments.
Higher enrollments are still associated with lower FX_Rates, but the GDP/Inflation values don't show a strong, direct correlation in this plot.
The clipped data suggests that while FX_Rates affect enrollment numbers, economic indicators might have more subtle or context-dependent effects.

4. Enrollment vs FX_Rate with Growth Indicators (GDP/Inflation Values) - No Outlier
Observation:
Without outliers, the distribution of growth indicators is more consistent across the enrollment spectrum.
The plot indicates that both FX_Rates and economic conditions (represented by GDP/Inflation) can influence enrollment, but the impact seems less extreme when outliers are removed.
The trend of decreasing enrollment with increasing FX_Rates is still evident, but the spread of economic conditions is more uniform.

General Analysis Across All Plots:
FX_Rate Impact: There is a clear inverse relationship between FX_Rates and student enrollment, where higher FX_Rates tend to correlate with lower enrollment numbers.
Effect of Tuition Fees: Tuition fees also seem to follow a similar pattern where higher fees are more common at lower enrollment levels, particularly when FX_Rates are high.
Growth Indicators: Economic conditions like GDP and inflation appear to play a role but are not the sole factors influencing enrollment trends. The clipped vs. no outlier data shows that extreme values can skew the interpretation, highlighting the importance of considering outliers in such analyses.
Overall, these plots reveal that student enrollment is sensitive to FX_Rates, tuition fees, and possibly to broader economic conditions, though the relationship is complex and may vary depending on the specific context or country.


#------CORRELATION TRSTS BEFORE ONE HOT ENCODING AND MODEL BUILDING - no outlier dataset will be used moving forward
```{r}
# Pearson Correlation Test
pearson_corr <- cor.test(final_combined_1422_no_outlier$FX_Rate, final_combined_1422_no_outlier$Enrollment, method = "pearson")

# Print Pearson correlation results
print("Pearson Correlation: FX on Enrollment Numbers")
print(pearson_corr)

# Pearson Correlation Test
pearson_corr2 <- cor.test(final_combined_1422_no_outlier$Value, final_combined_1422_no_outlier$Enrollment, method = "pearson")

# Print Pearson correlation results
print("Pearson Correlation: GDP Value on Enrollment Numbers")
print(pearson_corr2)

# Pearson Correlation Test
pearson_corr3 <- cor.test(final_combined_1422_no_outlier$Tuition_Av, final_combined_1422_no_outlier$Enrollment, method = "pearson")

# Print Pearson correlation results
print("Pearson Correlation: tuition fees on Enrollment Numbers")
print(pearson_corr3)

# Pearson Correlation Test
#final_combined_1422_no_outlier$Visa.Fee.... <- gsub("[^0-9]", "", final_combined_1422_no_outlier$Visa.Fee....)
final_combined_1422_no_outlier$Visa.Fee.... = as.numeric(final_combined_1422_no_outlier$Visa.Fee....)
pearson_corr4 <- cor.test(final_combined_1422_no_outlier$Visa.Fee...., final_combined_1422_no_outlier$Enrollment, method = "pearson")

# Print Pearson correlation results
print("Pearson Correlation: Visa fees on Enrollment Numbers")
print(pearson_corr4)

#summary(final_combined_1422_no_outlier$Visa.Fee....)
```
FX Rate and GDP Value: Both have very weak positive correlations with enrollment numbers, but these relationships are statistically significant, suggesting they are not due to chance.
Tuition Fees: There is an extremely weak positive correlation with enrollment numbers, but this is not statistically significant, indicating it could be due to random variation.
Visa Fees: A weak negative correlation exists between visa fees and enrollment numbers, and this relationship is statistically significant, suggesting a genuine link.


##---------------------------------------------------ONE HOT ENCODING--------------------------------------------

```{r}
# Update the list of categorical columns to include all necessary variables
categorical_columns <- c("HE.Provider", "Mode.Of..Study", "Country", 
                         "Visa.Code", "Growth.Indicator", "Visa.After.Graduation", 
                         "Visa.Work.Requirements.After.Graduation")

# Create a formula for model.matrix, encoding all categorical variables
formula <- as.formula(paste("~", paste(categorical_columns, collapse = " + "), "- 1"))

# Apply one-hot encoding using model.matrix()
one_hot_encoded_data <- model.matrix(formula, data = final_combined_1422_no_outlier)

# Convert the resulting matrix back to a data frame
one_hot_encoded_df <- as.data.frame(one_hot_encoded_data)

# Combine with the original dataset (excluding the original categorical columns)
final_combined_1422_encoded <- cbind(final_combined_1422_no_outlier[, !names(final_combined_1422_no_outlier) %in% categorical_columns], one_hot_encoded_df)

# View the resulting data frame with one-hot encoded columns
head(final_combined_1422_encoded)
```

##----------------------CORRELATION - ONE HOT ENCODING ------------------------------------------

```{r,echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
final_combined_1422_encoded$Visa.NHS.Charge....= as.numeric(final_combined_1422_encoded$Visa.NHS.Charge....)
final_combined_1422_encoded$Visa.Fee..Dependents = as.numeric(final_combined_1422_encoded$Visa.Fee..Dependents)
final_combined_1422_encoded$Visa.NHS.Charge.Depen = as.numeric(final_combined_1422_encoded$Visa.NHS.Charge.Depen)
# Calculate the correlation matrix using pearson
 corrplot(cor(final_combined_1422_encoded[,c("FX_Rate", "Enrollment", "Visa.Fee....","Visa.NHS.Charge....", "Value", "Tuition_Av", "Visa.Fee..Dependents", "Visa.NHS.Charge.Depen")]), type = 'upper', method = 'number')

# Calculate the correlation matrix using the Spearman method
cor_matrix <- cor(final_combined_1422_encoded[, c("FX_Rate", "Enrollment", "Visa.Fee....", "Visa.NHS.Charge....", "Value", "Tuition_Av", "Visa.Fee..Dependents", "Visa.NHS.Charge.Depen")], method = "spearman")

# Visualize the correlation matrix using corrplot
corrplot(cor_matrix, method = "number", type = "upper")
```

Summary---
FX Rate on Enrollment Numbers:

Correlation coefficient (r): 0.0501
P-value: 0.0003577
Confidence Interval: [0.0226, 0.0775]
Interpretation: There is a very weak positive correlation between the FX rate and enrollment numbers. The correlation is statistically significant (p < 0.001), meaning that even though the relationship is weak, it is unlikely to be due to random chance.
GDP Value on Enrollment Numbers:

Correlation coefficient (r): 0.0362
P-value: 0.009811
Confidence Interval: [0.0087, 0.0637]
Interpretation: There is a very weak positive correlation between GDP value and enrollment numbers. The correlation is statistically significant (p < 0.01), indicating that the observed correlation, although weak, is not likely due to chance.
Tuition Fees on Enrollment Numbers:

Correlation coefficient (r): 0.0229
P-value: 0.1035
Confidence Interval: [-0.0047, 0.0503]
Interpretation: There is an extremely weak positive correlation between tuition fees and enrollment numbers. However, this correlation is not statistically significant (p > 0.05), suggesting that the observed relationship could be due to random variation rather than an actual association.
Visa Fees on Enrollment Numbers:

Correlation coefficient (r): -0.0353
P-value: 0.01185
Confidence Interval: [-0.0628, -0.0078]
Interpretation: There is a very weak negative correlation between visa fees and enrollment numbers. The correlation is statistically significant (p < 0.05), indicating that the relationship, although weak, is likely to be real rather than due to chance.
Summary
FX Rate and GDP Value both show weak positive correlations with enrollment numbers, but these correlations are statistically significant.
Tuition Fees show an extremely weak and statistically insignificant positive correlation with enrollment numbers.
Visa Fees show a weak negative correlation with enrollment numbers, which is statistically significant.



##Identify which currencies have the strongest correlation with enrollment changes.
```{r}
#using encoded dataset
country_list <- c("Gambia", "Ghana", "India", "Japan", "Kenya", "Kuwait", "Libya", "Nigeria", "Oman", "Pakistan", "Tanzania", "Thailand", "Turkey", "Uganda")

#  vector to store correlations
correlations <- numeric(length(country_list))

# Calculate correlation for each country
for (i in seq_along(country_list)) {
  country <- country_list[i]
  country_enrollment <- final_combined_1422_encoded[[paste0("Country", country)]]
  correlations[i] <- cor(final_combined_1422_encoded$FX_Rate, country_enrollment, method = "pearson")
}

# Create named vector for easy reference
names(correlations) <- country_list

# Sort by absolute value to find the strongest correlations
sorted_correlations <- sort(abs(correlations), decreasing = TRUE)

# Display the sorted correlations
sorted_correlations
```
```{r}
#visualise it

barplot(sorted_correlations, 
        main = "Correlation Between FX_Rate and Enrollment by Country", 
        las = 2, 
        col = "lightblue", 
        ylab = "Absolute Correlation Coefficient",
        names.arg = names(sorted_correlations))
```

## Correlation and Multivariate Analyis to explore how other variables affect enrollment
--------------------------------------------------------------------------

```{r,echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
library(knitr)
library(kableExtra)
# Ensure Fiscal_Year is numeric
final_combined_1422_encoded <- final_combined_1422_encoded %>%
  mutate(Fiscal_Year = as.numeric(as.character(Fiscal_Year)))

# Replace spaces and periods with underscores, remove commas
colnames(final_combined_1422_encoded) <- gsub("[[:space:].,]", "_", colnames(final_combined_1422_encoded))

# Optionally, you can also convert everything to lowercase for consistency
colnames(final_combined_1422_encoded) <- tolower(colnames(final_combined_1422_encoded))

colnames(final_combined_1422_encoded)

# Calculate correlation matrix including new variables
cor_matrix1 <- cor(final_combined_1422_encoded[, c(
"enrollment", 
  "fx_rate", 
  "visa_fee____", 
  "tuition_av", 
  "value",
  "visa_codetier_4_(general)", 
  "visa_codetier_4_+_graduate_route", 
  "visa_codetier_4_pilot",
  "visa_after_graduationtier_2_only", 
  "visa_after_graduationyes", 
  "visa_work_requirements_after_graduationgraduate_route_for_2_years",
  "visa_work_requirements_after_graduationnone"
)], method = 'spearman')


# Convert the correlation matrix to a data frame
cor_matrix1_df <- as.data.frame(as.table(cor_matrix1))


print(cor_matrix1_df)

# Render the table in a nicely formatted way
cor_matrix1_df %>%
  kable("html", caption = "Correlation Matrix") %>%
  kable_styling(full_width = F)

#
# Re-run your model
lm_model <- lm(enrollment ~ fx_rate + `visa_fee____` + `tuition_av` + value, data = final_combined_1422_encoded)
summary(lm_model)

# Feature Importance using Random Forest
rf_model2 <- randomForest(enrollment ~ fx_rate + visa_fee____ + tuition_av + value, data = final_combined_1422_encoded)
importance(rf_model2)
```

```{r}
colnames(final_combined_1422_encoded)
```


#------------------------------ -------------------------------------------------- Moving to Statistical Tests, Correlation and Non-model Based approaches -------------------------------------------------------------------


#Objective 1 -- 1. Quantifying the Relationship of FX Rate by Enrollment
Objective: Understand how exchange rate fluctuations affect international student enrollment.
```{r}
#General
pearson_corr <- cor(final_combined_1422_encoded$fx_rate, final_combined_1422_encoded$enrollment, method = "pearson")
print(pearson_corr)
```

By Country
```{r,echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
#Same thing but by country
# List of country columns (assumed that each country column has enrollment data)
country_columns <- c("countryindia", "countryjapan", "countrykenya", 
                     "countrykuwait", "countrylibya", "countrynigeria", 
                     "countryoman", "countrypakistan", "countrytanzania","countrygambia","countryghana" ,
                     "countrythailand", "countryturkey", "countryuganda")

# Initialize data frame to store correlation results
correlation_results <- data.frame(Country = character(), Pearson = numeric(), Spearman = numeric(), stringsAsFactors = FALSE)

# Loop through each country to calculate correlations
for (country in country_columns) {
  # Assuming 'FX_Rate' is the same for all countries, otherwise adjust accordingly
  pearson_corr <- cor(final_combined_1422_encoded[[country]], final_combined_1422_encoded$fx_rate, method = "pearson", use = "complete.obs")
  spearman_corr <- cor(final_combined_1422_encoded[[country]], final_combined_1422_encoded$fx_rate, method = "spearman", use = "complete.obs")
  
  # Append results to the data frame
  correlation_results <- rbind(correlation_results, data.frame(Country = country, Pearson = pearson_corr, Spearman = spearman_corr))
}

# Print the correlation results
print(correlation_results)
```


```{r,echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
#---- visualise

# Loop through each country to calculate correlations
for (country in country_columns) {
  # Convert the country column to numeric (handling any non-numeric data)
  country_data <- as.numeric(as.character(final_combined_1422_encoded[[country]]))
  
  # Skip if the conversion leads to NA (i.e., non-numeric column)
  if (all(is.na(country_data))) {
    next
  }
  
  # Calculate Pearson and Spearman correlations
  pearson_corr1 <- cor(country_data, final_combined_1422_encoded$fx_rate, method = "pearson", use = "complete.obs")
  spearman_corr1 <- cor(country_data, final_combined_1422_encoded$fx_rate, method = "spearman", use = "complete.obs")
  
  # Append results to the data frame
  correlation_results1 <- rbind(correlation_results, data.frame(Country = country, Pearson = pearson_corr1, Spearman = spearman_corr1))
}

# Reshape the data for plotting
correlation_results_long <- melt(correlation_results, id.vars = "Country", variable.name = "Correlation_Type", value.name = "Correlation")

# Plot the correlations using ggplot2
ggplot(correlation_results_long, aes(x = Country, y = Correlation, fill = Correlation_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +  # Flip coordinates for better readability
  labs(title = "Correlation between FX Rate and Enrollment by Country", x = "Country", y = "Correlation") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))  # Adjust text size for readability
```

By University
```{r,echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
# List of university columns (assuming each university column has enrollment data)
university_columns <- c(
  "he_providerabertay_university", 
  "he_provideredinburgh_napier_university", 
  "he_providerglasgow_caledonian_university", 
  "he_providerglasgow_school_of_art", 
  "he_providerheriot_watt_university", 
  "he_providerqueen_margaret_university__edinburgh", 
  "he_providerrobert_gordon_university", 
  "he_providerroyal_conservatoire_of_scotland", 
  "he_providersruc", 
  "he_providerthe_university_of_aberdeen", 
  "he_providerthe_university_of_dundee", 
  "he_providerthe_university_of_edinburgh", 
  "he_providerthe_university_of_glasgow", 
  "he_providerthe_university_of_st__andrews", 
  "he_providerthe_university_of_stirling", 
  "he_providerthe_university_of_strathclyde", 
  "he_providerthe_university_of_the_west_of_scotland", 
  "he_provideruniversity_of_the_highlands_and_islands"
)

# Initialize data frame to store correlation results
correlation_results <- data.frame(University = character(), Pearson = numeric(), Spearman = numeric(), stringsAsFactors = FALSE)

# Loop through each university to calculate correlations
for (university in university_columns) {
  # Convert the university column to numeric (handling any non-numeric data)
  university_data <- as.numeric(as.character(final_combined_1422_encoded[[university]]))
  
  # Skip if the conversion leads to NA (i.e., non-numeric column)
  if (all(is.na(university_data))) {
    next
  }
  
  # Calculate Pearson and Spearman correlations
  pearson_corr <- cor(university_data, final_combined_1422_encoded$fx_rate, method = "pearson", use = "complete.obs")
  spearman_corr <- cor(university_data, final_combined_1422_encoded$fx_rate, method = "spearman", use = "complete.obs")
  
  # Append results to the data frame
  correlation_results <- rbind(correlation_results, data.frame(University = university, Pearson = pearson_corr, Spearman = spearman_corr))
}

# Reshape the data for plotting
correlation_results_long <- melt(correlation_results, id.vars = "University", variable.name = "Correlation_Type", value.name = "Correlation")

# Plot the correlations using ggplot2
ggplot(correlation_results_long, aes(x = University, y = Correlation, fill = Correlation_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +  # Flip coordinates for better readability
  labs(title = "Correlation between FX Rate and Enrollment by University", x = "University", y = "Correlation") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))  # Adjust text size for readability
```

Granger Causality Test: Test if changes in exchange rates can predict changes in enrollment
```{r}
library(lmtest)
granger_test <- grangertest(enrollment ~ fx_rate, order = 1, data = final_combined_1422_encoded)
print(granger_test)
```


```{r}
#General
pearson_corr_general <- cor(final_combined_1422_encoded$fx_rate, final_combined_1422_encoded$enrollment, method = "pearson")
print(pearson_corr_general)

# By Visa Code
# List of visa code columns (assumed that each visa code column has enrollment data)
visa_code_columns <- c("visa_codetier_4_(general)", "visa_codetier_4_+_graduate_route", "visa_codetier_4_pilot")

# Initialize data frame to store correlation results
correlation_results_visa <- data.frame(Visa_Code = character(), Pearson_FX_Rate = numeric(), Spearman_FX_Rate = numeric(), 
                                       Pearson_Enrollment = numeric(), Spearman_Enrollment = numeric(), 
                                       Pearson_Tuition_Fees = numeric(), Spearman_Tuition_Fees = numeric(),
                                       Pearson_Value = numeric(), Spearman_Value = numeric(),
                                       Pearson_Visa_Fee = numeric(), Spearman_Visa_Fee = numeric(),
                                       stringsAsFactors = FALSE)

# Loop through each visa code to calculate correlations
for (visa_code in visa_code_columns) {
  # Calculate Pearson and Spearman correlations for fx_rate, enrollment, tuition_fees, value, and visa_fee
  pearson_corr_fx_rate <- cor(final_combined_1422_encoded[[visa_code]], final_combined_1422_encoded$fx_rate, method = "pearson", use = "complete.obs")
  spearman_corr_fx_rate <- cor(final_combined_1422_encoded[[visa_code]], final_combined_1422_encoded$fx_rate, method = "spearman", use = "complete.obs")

  pearson_corr_enrollment <- cor(final_combined_1422_encoded[[visa_code]], final_combined_1422_encoded$enrollment, method = "pearson", use = "complete.obs")
  spearman_corr_enrollment <- cor(final_combined_1422_encoded[[visa_code]], final_combined_1422_encoded$enrollment, method = "spearman", use = "complete.obs")

  pearson_corr_tuition_fees <- cor(final_combined_1422_encoded[[visa_code]], final_combined_1422_encoded$tuition_av, method = "pearson", use = "complete.obs")
  spearman_corr_tuition_fees <- cor(final_combined_1422_encoded[[visa_code]], final_combined_1422_encoded$tuition_av, method = "spearman", use = "complete.obs")

  pearson_corr_value <- cor(final_combined_1422_encoded[[visa_code]], final_combined_1422_encoded$value, method = "pearson", use = "complete.obs")
  spearman_corr_value <- cor(final_combined_1422_encoded[[visa_code]], final_combined_1422_encoded$value, method = "spearman", use = "complete.obs")

  pearson_corr_visa_fee <- cor(final_combined_1422_encoded[[visa_code]], final_combined_1422_encoded$visa_fee____, method = "pearson", use = "complete.obs")
  spearman_corr_visa_fee <- cor(final_combined_1422_encoded[[visa_code]], final_combined_1422_encoded$visa_fee____, method = "spearman", use = "complete.obs")

  # Append results to the data frame
  correlation_results_visa <- rbind(correlation_results_visa, 
                                    data.frame(Visa_Code = visa_code, 
                                               Pearson_FX_Rate = pearson_corr_fx_rate, Spearman_FX_Rate = spearman_corr_fx_rate,
                                               Pearson_Enrollment = pearson_corr_enrollment, Spearman_Enrollment = spearman_corr_enrollment,
                                               Pearson_Tuition_Fees = pearson_corr_tuition_fees, Spearman_Tuition_Fees = spearman_corr_tuition_fees,
                                               Pearson_Value = pearson_corr_value, Spearman_Value = spearman_corr_value,
                                               Pearson_Visa_Fee = pearson_corr_visa_fee, Spearman_Visa_Fee = spearman_corr_visa_fee))
}

# Print the correlation results
print(correlation_results_visa)
```


##Objective 3. Developing Predictive Insights
Objective: exploratory analysis that might hint at predictive relationships.

```{r,echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
#Time Series Analysis
# Convert your enrollment data to a time series object
enrollment_ts <- ts(final_combined_1422_encoded$enrollment, frequency = 12)  # Adjust frequency based on your data

# Apply STL decomposition
decomposition_stl <- stl(enrollment_ts, s.window = "periodic")

# Plot the decomposition
plot(decomposition_stl)
```


#3. Developing Predictive Insights
Objective: Develop insights to forecast enrollment based on currency fluctuations.

```{r}
library(forecast)
arima_model <- auto.arima(final_combined_1422_encoded$enrollment)
print(summary(arima_model))
```
Interpretation of the ARIMA(5,0,4) Model Results:
Model Overview
ARIMA(5,0,4): Uses 5 autoregressive terms, no differencing, and 4 moving average terms.
Non-zero Mean: The series has a non-zero mean of 19.3985.
Model Fit
Residuals: The model has a moderate level of error, as indicated by the RMSE and MAE.
Autocorrelation: Residuals are not autocorrelated, suggesting the model captures the serial correlation in the data.
Model Selection: The AIC and BIC values suggest a reasonable fit, but overfitting might be a concern due to the number of parameters.
Potential Issues
MAPE: The infinite MAPE indicates potential issues with zero or near-zero values in the data.
Model Complexity: The model's complexity might contribute to overfitting.
Recommendations
Data Preprocessing: Address any zero or near-zero values.
Model Validation: Evaluate the model on a separate test set.
Alternative Models: Consider simpler ARIMA configurations or other time series models.
Stationarity Check: Assess if differencing is necessary to make the data stationary.


#4. Investigating Student Sensitivity
Objective: Determine if certain student demographics are more sensitive to currency fluctuations.
```{r,echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
# Initialise data frame to store correlation results
segmented_corr <- data.frame(Country = character(), Pearson_Corr = numeric(), Spearman_Corr = numeric(), stringsAsFactors = FALSE)

# Loop through each country column to calculate correlations
for (country in country_columns) {
  pearson_corr <- cor(final_combined_1422_encoded[[country]], final_combined_1422_encoded$fx_rate, method = "pearson", use = "complete.obs")
  spearman_corr <- cor(final_combined_1422_encoded[[country]], final_combined_1422_encoded$fx_rate, method = "spearman", use = "complete.obs")
  
  # Append results to the data frame
  segmented_corr <- rbind(segmented_corr, data.frame(Country = country, Pearson_Corr = pearson_corr, Spearman_Corr = spearman_corr))
}

# Print the correlation results
print(segmented_corr)

# Plot the results
# Pearson Correlation Plot
ggplot(segmented_corr, aes(x = Country, y = Pearson_Corr)) +
  geom_bar(stat = "identity", fill = "blue", alpha = 0.7) +
  geom_text(aes(label = round(Pearson_Corr, 2)), vjust = -0.5, size = 6) +  # Increase label text size
  labs(title = "Pearson Correlation Between FX Rate and Enrollment by Country", 
       x = "Country", y = "Pearson Correlation Coefficient") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20, face = "bold"),  # Increase title font size
    axis.title.x = element_text(size = 15),               # Increase x-axis title font size
    axis.title.y = element_text(size = 15),               # Increase y-axis title font size
    axis.text.x = element_text(size = 12),                # Increase x-axis labels font size
    axis.text.y = element_text(size = 12)                 # Increase y-axis labels font size
  ) +
  coord_flip()  # Flip axis for better readability

# Spearman Correlation Plot
ggplot(segmented_corr, aes(x = Country, y = Spearman_Corr)) +
  geom_bar(stat = "identity", fill = "green", alpha = 0.7) +
  geom_text(aes(label = round(Spearman_Corr, 2)), vjust = -0.5, size = 6) +  # Increase label text size
  labs(title = "Spearman Correlation Between FX Rate and Enrollment by Country", 
       x = "Country", y = "Spearman Correlation Coefficient") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20, face = "bold"),  # Increase title font size
    axis.title.x = element_text(size = 15),               # Increase x-axis title font size
    axis.title.y = element_text(size = 15),               # Increase y-axis title font size
    axis.text.x = element_text(size = 12),                # Increase x-axis labels font size
    axis.text.y = element_text(size = 12)                 # Increase y-axis labels font size
  ) +
  coord_flip()  # Flip axis for better readability

#--------STOP---------------------#
```


Cluster Analyis
```{r}
# You can use clustering methods like K-means to identify groups
library(cluster)
set.seed(123)
data_scaled <- scale(final_combined_1422_encoded[, c("fx_rate", "mode_of__studypart-time", "value", "enrollment")])  # Scale data for clustering
kmeans_result <- kmeans(data_scaled, centers = 5)  # Adjust centers as needed
final_combined_1422_encoded$cluster <- kmeans_result$cluster
print(table(final_combined_1422_encoded$cluster))


#----visualisations
# Scatter plot using two variables (e.g., fx_rate vs. enrollment)
ggplot(final_combined_1422_encoded, aes(x = fx_rate, y = enrollment, color = as.factor(cluster))) +
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "K-means Clustering: fx_rate vs Enrollment",
       x = "FX Rate", y = "Enrollment",
       color = "Cluster") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20, face = "bold"),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  )

#======
library(factoextra)

# PCA plot to visualise clusters
fviz_cluster(kmeans_result, data = data_scaled,
             geom = "point", 
             ellipse.type = "convex",
             palette = "jco",
             ggtheme = theme_minimal()) +
  labs(title = "K-means Clustering Visualization using PCA",
       x = "Principal Component 1",
       y = "Principal Component 2")
```

`
```{r}
pca_result <- prcomp(data_scaled, center = TRUE, scale. = TRUE)

# Print summary of PCA
summary(pca_result)
```
```{r}
loadings <- pca_result$rotation

# View the loadings for PC1 and PC2
print(loadings[, 1:4])  # For PC1 and PC2
```
```{r,echo=FALSE, warning=FALSE, message=FALSE, fig.width=14, fig.height=8}
# Biplot showing PCA loadings and scores
biplot(pca_result, scale = 0)
```




####------------------------------code has errors from here ---------------------------------------------------------------------------------------
#----------- FEATURE ENGINEERING---------------------------



1. Calculate rolling averages of exchange rates and economic indicators to smooth out short-term fluctuations and focus on longer-term trends.
```{r}
final_combined_1422_encoded <- final_combined_1422_encoded %>%
  mutate(
    FX_Rate_Rolling_Interaction = fx_rate * rolling_average,
    Tuition_Rolling_Interaction = tuition_av * rolling_average,
    Value_Rolling_Interaction = value * rolling_average
  )
```


2.  Create interaction terms between currency exchange rates and economic indicators to capture potential combined effects. 
```{r}
final_combined_1422_encoded <- final_combined_1422_encoded %>%
  mutate(
    FX_Rate_Rolling_Interaction = fx_rate * rolling_average,
    Tuition_Rolling_Interaction = tuition_av * rolling_average,
    Value_Rolling_Interaction = value * rolling_average
  )
```

```{r}
# Remove rows with NA values
final_combined_1422_encoded_ <- na.omit(final_combined_1422_encoded)
# Remove rows with NA values in specific columns
final_combined_1422_encoded <- final_combined_1422_encoded %>%
  drop_na(fx_rate, visa_fee____,tuition_av, value)

```

#feature selection using random forest (  trying to understand the feature importance)
```{r}
# Set a seed for reproducibility
set.seed(123)

# Check for missing values in the dataset
colSums(is.na(final_combined_1422_encoded))

# Remove rows with missing values
final_combined_1422_encoded <- na.omit(final_combined_1422_encoded)

# Example of mean imputation
final_combined_1422_encoded[is.na(final_combined_1422_encoded)] <- 
    lapply(final_combined_1422_encoded, function(x) ifelse(is.na(x), mean(x, na.rm = TRUE), x))

# Split the data into training (80%) and test (20%) sets
train_index <- createDataPartition(final_combined_1422_encoded$enrollment, p = 0.8, list = FALSE)
train_data <- final_combined_1422_encoded[train_index, ]
test_data <- final_combined_1422_encoded[-train_index, ]

if("he_providerheriot-watt_university" %in% colnames(train_data)) {
    print("Column exists.")
} else {
    print("Column does not exist in train_data.")
}

#any(is.na(train_data$he_providerheriot-watt_university))


colnames(train_data) <- gsub("-", "_", colnames(train_data))
if("he_providerheriot_watt_university" %in% colnames(train_data)) {
    print("Renamed column exists.")
} else {
    print("Renamed column does not exist in train_data.")
}

colnames(train_data)[colnames(train_data) == "he_providerheriot-watt_university"] <- "he_providerheriot_watt_university"

any(is.na(train_data$he_providerheriot_watt_university))

colnames(train_data) <- gsub("\\(", "_", colnames(train_data))
colnames(train_data) <- gsub("\\)", "_", colnames(train_data))
colnames(train_data) <- gsub("\\+", "_", colnames(train_data))

# Double check if the columns are correctly renamed
if("visa_codetier_4__general_" %in% colnames(train_data)) {
    print("Renamed 'visa_codetier_4_(general)' exists.")
} else {
    print("Renamed 'visa_codetier_4_(general)' does not exist in train_data.")
}

if("visa_codetier_4___graduate_route" %in% colnames(train_data)) {
    print("Renamed 'visa_codetier_4_+_graduate_route' exists.")
} else {
    print("Renamed 'visa_codetier_4_+_graduate_route' does not exist in train_data.")
}

# Train a Random Forest model on the training data
rf_model_fs <- randomForest(enrollment ~ ., data = train_data, importance = TRUE)

# View the importance of each feature
importance(rf_model_fs)
print(rf_model_fs)
```


```{r}
# Select top features based on importance
colnames(train_data) <- tolower(colnames(train_data))
colnames(train_data) <- gsub("\\s+", "_", colnames(train_data))

selected_features <- c("mode_of__studypart_time" , "rolling_average", "tuition_rolling_interaction", "fx_rate", "fx_rate_rolling_interaction", "Value_rolling_interaction ","value") # Replace with top features

### BEUGGING~~~
#missing_features <- selected_features[!selected_features %in% colnames(train_data)]
#if (length(missing_features) > 0) {
    #print("The following selected features are missing from train_data:")
   # print(missing_features)
#} else {
 #   print("All selected features are present in train_data.")
#}

# Remove any leading or trailing whitespace from column names
#colnames(train_data) <- trimws(colnames(train_data))

# Check for the presence of the 'Value_Rolling_Interaction' column again
if ("Value_Rolling_Interaction" %in% colnames(train_data)) {
    print("Column 'Value_Rolling_Interaction' exists in train_data.")
} else {
    print("Column 'Value_Rolling_Interaction' does NOT exist in train_data.")
}

# Print all column names to inspect them
print(colnames(train_data))

colnames(train_data) <- tolower(colnames(train_data))
colnames(train_data) <- gsub("\\s+", "_", colnames(train_data))

setdiff(selected_features, colnames(train_data))

print(colnames(train_data))

# Train the model using the selected features
model_selected <- lm(enrollment ~ ., data = train_data[, c("enrollment", selected_features)])
summary(model_selected)
```


```{r}
# Predict on the test data using the trained model
predictions_fs <- predict(model_selected, newdata = test_data[, selected_features])

# Calculate performance metrics
mse <- mean((test_data$Enrollment - predictions)^2)
rmse <- sqrt(mse)
r_squared <- cor(test_data$Enrollment, predictions)^2

# Print the performance metrics
print(paste("RMSE: ", rmse))
print(paste("R-squared: ", r_squared))
```























